Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-07-15 thirdPaper/draft/typeConeAssociahedra.tex
Generalizes --> Generalizes to
3 typos

We now give a more detailed overview of our main results and the structure of the article. In particular, some of the geometric realizations in~\cref{part:geometry} depend on algebraic results whose proof is delayed until~\cref{part:algebra}. The goal of this section is to clarify their logical dependencies, summarized in the schematic diagram depicted at the end.

@@

The two main ingredients are similar to those for cluster algebras.
As a first step, we obtain the following consequence of \cref{prop:exchangeablePairsNKC} (ii), which is proved in \cref{sec:Kbproj} by representation theory.

@@

\newtheorem*{thm:extricats}{\cref{thm:extricats}}
\begin{thm:extricats}
The extriangulated category~$\cat$ has only finitely many ismorphism classes of indecomposable objects if and only if the set~$L_\cat$ generates the kernel of the canonical projection ${\b{g}:\Ksp(\cat)\to\kzero{\cat}}$.
In that case~$L_\cat$ is a basis of the kernel of~$\b{g}$ and any~$x\in\ker(\b{g})$ decomposes~as
\[
 x = \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]
\end{thm:extricats}

@@

\vspace{.5cm}
\tikzset{
  double arrow/.style={
    -latex, line width=3pt, black, % first arrow
    postaction={draw,line width=2pt, white, shorten <= 1,shorten >= 2}, % second arrow
  }
}
\centerline{
\begin{tikzpicture}[every node/.style={inner sep=6, outer sep=8}]
	\node[diamond,aspect=3, text width=5cm, minimum height=1cm, align=center, draw=blue, fill=blue!10, inner sep=-2] (I1) at (0,1.5) {\ref{part:geometry}.\ref{sec:typeCone} -- Simplicial type cone strategy};
	\node[rectangle, draw=black, fill=black!10, align=center] (I21) at (-6,0) {\ref{part:geometry}.\ref{subsec:typeConeAsso} -- Toy example \\ Classical associahedra};
	\node[rectangle, draw=orange, fill=orange!10, align=center, text width=5cm] (I22) at (-3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeCA} -- Cluster algebras \\ \& generalized associahedra};
	\node[rectangle, draw=green, fill=green!10, align=center, text width=5cm] (I23) at (3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeNKC} -- Gentle algebras \\ \& generalized accordiohedra};
	\node[rectangle, draw=red, fill=red!10, align=center, text width=5.4cm] (I24) at (7,0) {\ref{part:geometry}.\ref{subsec:typeConeGA} -- Graphical nested complexes \\ \& graph associahedra};
	\node[ellipse, draw=orange, fill=orange!10, align=center, text width=4.9cm, inner sep=2, minimum height=1cm] (II1) at (-4.5,-4.2) {\ref{part:algebra}.\ref{sec:clusterCategories} -- Cluster categories};
	\node[ellipse, draw=green, fill=green!10, align=center, text width=4.9cm, inner sep=0, minimum height=1cm] (II2) at (4.5,-4.2) {\ref{part:algebra}.\ref{sec:extricats} -- Extriangulated categories};
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.180) to [bend right, looseness=1.1] node[label={[shift={(.6,-1.5)}]applied to}]{} (I21.90);
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.220) to [bend right, looseness=.6] node[label={[shift={(.6,-1.5)}]applied to}]{} (I22.90);
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.320) to [bend left, looseness=.6] node[label={[shift={(-.6,-1.5)}]applied to}]{} (I23.90);
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=10}, draw=blue, fill=blue!10] (I1.0) to [bend left, looseness=1.1] (I24.90);
	\draw[decorate, decoration={markings, mark=at position 0.5 with {\draw [red, ultra thick,-] ++ (-.3,-.3)  -- ( .3, .3); \draw [red, ultra thick,-] ++ (-.3, .3)  -- ( .3,-.3);}}] (I1.0) to [bend left, looseness=1.1] node[label={[shift={(-.4,-1.7)}]not applicable to}]{} (I24.90);
	\draw[thick, double distance=2pt, -implies] (I22.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I22.270)+(0,.15);
	\draw[thick, double distance=2pt, -implies] (I23.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I23.270)+(0,.15);
	\draw[thick, ->] (II1.0) to node[semithick, above, outer sep=0]{generalizes} (II2.180);
	\draw let \p{top}=(I1.90), \p{left}=(I21.180), \p{bottom}=(I22.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\draw let \p{top}=([shift={(0,.2)}]II1.90), \p{left}=(I21.180), \p{bottom}=(II1.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\path let \p{top}=(I1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-6.15, \y{top}) {\ref{part:geometry} -- Discrete geometry};
	\path let \p{top}=([shift={(0,.2)}]II1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-5.8, \y{top}) {\ref{part:algebra} -- Representation theory};
\end{tikzpicture}
}

\centerline{
\begin{tikzpicture}[every node/.style={inner sep=6, outer sep=8}]
 	\node[cloud, cloud puffs=25, cloud ignores aspect, text width=5cm, minimum height=1cm, align=center, draw=blue, fill=blue!10, inner sep=3, outer sep=5] (I1) at (0,1.5) {\ref{part:geometry}.\ref{sec:typeCone} -- Simplicial type cone strategy};
	\node[rectangle, draw=black, fill=black!10, align=center] (I21) at (-6,0) {\ref{part:geometry}.\ref{subsec:typeConeAsso} -- Toy example \\ Classical associahedra};
	\node[rectangle, draw=orange, fill=orange!10, align=center, text width=5cm] (I22) at (-3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeCA} -- Cluster algebras \\ \& generalized associahedra};
	\node[rectangle, draw=green, fill=green!10, align=center, text width=5cm] (I23) at (3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeNKC} -- Gentle algebras \\ \& generalized accordiohedra};
	\node[rectangle, draw=red, fill=red!10, align=center, text width=5.4cm] (I24) at (7,0) {\ref{part:geometry}.\ref{subsec:typeConeGA} -- Graphical nested complexes \\ \& graph associahedra};
	\node[ellipse, draw=orange, fill=orange!10, align=center, text width=4.9cm, inner sep=2, minimum height=1cm] (II1) at (-4.5,-4.2) {\ref{part:algebra}.\ref{sec:clusterCategories} -- Cluster categories};
	\node[ellipse, draw=green, fill=green!10, align=center, text width=4.9cm, inner sep=0, minimum height=1cm] (II2) at (4.5,-4.2) {\ref{part:algebra}.\ref{sec:extricats} -- Extriangulated categories};
 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.180) to [bend right, looseness=1.1] node[label={[shift={(.6,-1.5)}]applied to}]{} (I21.90);
 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.240) to [bend right, looseness=.6] node[label={[shift={(.6,-1.5)}]applied to}]{} (I22.90);
 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.300) to [bend left, looseness=.6] node[label={[shift={(-.6,-1.5)}]applied to}]{} (I23.90);
 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=10}, draw=blue, fill=blue!10] (I1.0) to [bend left, looseness=1.1] (I24.90);
 	\draw[decorate, decoration={markings, mark=at position 0.5 with {\draw [red, ultra thick,-] ++ (-.3,-.3)  -- ( .3, .3); \draw [red, ultra thick,-] ++ (-.3, .3)  -- ( .3,-.3);}}] (I1.0) to [bend left, looseness=1.1] node[label={[shift={(-.4,-1.7)}]not applicable to}]{} (I24.90);
	\draw[thick, double distance=2pt, -implies] (I22.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I22.270)+(0,.15);
	\draw[thick, double distance=2pt, -implies] (I23.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I23.270)+(0,.15);
	\draw[thick, ->] (II1.0) to node[semithick, above, outer sep=0]{generalizes} (II2.180);
	\draw let \p{top}=(I1.90), \p{left}=(I21.180), \p{bottom}=(I22.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\draw let \p{top}=([shift={(0,.2)}]II1.90), \p{left}=(I21.180), \p{bottom}=(II1.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\path let \p{top}=(I1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-6.15, \y{top}) {\ref{part:geometry} -- Discrete geometry};
	\path let \p{top}=([shift={(0,.2)}]II1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-5.8, \y{top}) {\ref{part:algebra} -- Representation theory};
\end{tikzpicture}
}

>>>>

We now give a more detailed overview of our main results and the structure of the article. In particular, some of the geometric realizations in~\cref{part:geometry} depend on algebraic results whose proves are delayed until~\cref{part:algebra}. The goal of this section is to clarify their logical dependencies, summarized in the schematic diagram depicted at the end.

@@

The two main ingredients are similar to those for cluster algebras.
As a first step, we obtain the following consequence of \cref{prop:exchangeablePairsNKC} (ii), which is proved in \cref{sec:Kbproj} by representation-theoretic methods.

@@

\newtheorem*{thm:extricats}{\cref{thm:extricats}}
\begin{thm:extricats}
The extriangulated category~$\cat$ has only finitely many isomorphism classes of indecomposable objects if and only if the set~$L_\cat$ generates the kernel of the canonical projection ${\b{g}:\Ksp(\cat)\to\kzero{\cat}}$.
In that case~$L_\cat$ is a basis of the kernel of~$\b{g}$ and any~$x\in\ker(\b{g})$ decomposes~as
\[
 x = \sum_{A \in \ind(\cat) \ssm \add(\susp T)} \frac{\langle x, A \rangle}{\langle \ell_A, A \rangle} \ell_A.
\]
\end{thm:extricats}

@@

\vspace{.5cm}
\tikzset{
  double arrow/.style={
    -latex, line width=3pt, black, % first arrow
    postaction={draw,line width=2pt, white, shorten <= 1,shorten >= 2}, % second arrow
  }
}
\centerline{
\begin{tikzpicture}[every node/.style={inner sep=6, outer sep=8}]
	\node[diamond,aspect=3, text width=5cm, minimum height=1cm, align=center, draw=blue, fill=blue!10, inner sep=-2] (I1) at (0,1.5) {\ref{part:geometry}.\ref{sec:typeCone} -- Simplicial type cone strategy};
	\node[rectangle, draw=black, fill=black!10, align=center] (I21) at (-6,0) {\ref{part:geometry}.\ref{subsec:typeConeAsso} -- Toy example \\ Classical associahedra};
	\node[rectangle, draw=orange, fill=orange!10, align=center, text width=5cm] (I22) at (-3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeCA} -- Cluster algebras \\ \& generalized associahedra};
	\node[rectangle, draw=green, fill=green!10, align=center, text width=5cm] (I23) at (3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeNKC} -- Gentle algebras \\ \& generalized accordiohedra};
	\node[rectangle, draw=red, fill=red!10, align=center, text width=5.4cm] (I24) at (7,0) {\ref{part:geometry}.\ref{subsec:typeConeGA} -- Graphical nested complexes \\ \& graph associahedra};
	\node[ellipse, draw=orange, fill=orange!10, align=center, text width=4.9cm, inner sep=2, minimum height=1cm] (II1) at (-4.5,-4.2) {\ref{part:algebra}.\ref{sec:clusterCategories} -- Cluster categories};
	\node[ellipse, draw=green, fill=green!10, align=center, text width=4.9cm, inner sep=0, minimum height=1cm] (II2) at (4.5,-4.2) {\ref{part:algebra}.\ref{sec:extricats} -- Extriangulated categories};
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.180) to [bend right, looseness=1.1] node[label={[shift={(.6,-1.5)}]applied to}]{} (I21.90);
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.220) to [bend right, looseness=.6] node[label={[shift={(.6,-1.5)}]applied to}]{} (I22.90);
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.320) to [bend left, looseness=.6] node[label={[shift={(-.6,-1.5)}]applied to}]{} (I23.90);
	\draw[decorate, decoration={shape backgrounds, shape=dart, shape size=2mm, shape evenly spread=10}, draw=blue, fill=blue!10] (I1.0) to [bend left, looseness=1.1] (I24.90);
	\draw[decorate, decoration={markings, mark=at position 0.5 with {\draw [red, ultra thick,-] ++ (-.3,-.3)  -- ( .3, .3); \draw [red, ultra thick,-] ++ (-.3, .3)  -- ( .3,-.3);}}] (I1.0) to [bend left, looseness=1.1] node[label={[shift={(-.4,-1.7)}]not applicable to}]{} (I24.90);
	\draw[thick, double distance=2pt, -implies] (I22.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I22.270)+(0,.15);
	\draw[thick, double distance=2pt, -implies] (I23.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I23.270)+(0,.15);
	\draw[thick, ->] (II1.0) to node[semithick, above, outer sep=0]{generalizes to} (II2.180);
	\draw let \p{top}=(I1.90), \p{left}=(I21.180), \p{bottom}=(I22.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\draw let \p{top}=([shift={(0,.2)}]II1.90), \p{left}=(I21.180), \p{bottom}=(II1.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\path let \p{top}=(I1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-6.15, \y{top}) {\ref{part:geometry} -- Discrete geometry};
	\path let \p{top}=([shift={(0,.2)}]II1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-5.8, \y{top}) {\ref{part:algebra} -- Representation theory};
\end{tikzpicture}
}

\centerline{
\begin{tikzpicture}[every node/.style={inner sep=6, outer sep=8}]
 	\node[cloud, cloud puffs=25, cloud ignores aspect, text width=5cm, minimum height=1cm, align=center, draw=blue, fill=blue!10, inner sep=3, outer sep=5] (I1) at (0,1.5) {\ref{part:geometry}.\ref{sec:typeCone} -- Simplicial type cone strategy};
	\node[rectangle, draw=black, fill=black!10, align=center] (I21) at (-6,0) {\ref{part:geometry}.\ref{subsec:typeConeAsso} -- Toy example \\ Classical associahedra};
	\node[rectangle, draw=orange, fill=orange!10, align=center, text width=5cm] (I22) at (-3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeCA} -- Cluster algebras \\ \& generalized associahedra};
	\node[rectangle, draw=green, fill=green!10, align=center, text width=5cm] (I23) at (3,-1.5) {\ref{part:geometry}.\ref{subsec:typeConeNKC} -- Gentle algebras \\ \& generalized accordiohedra};
	\node[rectangle, draw=red, fill=red!10, align=center, text width=5.4cm] (I24) at (7,0) {\ref{part:geometry}.\ref{subsec:typeConeGA} -- Graphical nested complexes \\ \& graph associahedra};
	\node[ellipse, draw=orange, fill=orange!10, align=center, text width=4.9cm, inner sep=2, minimum height=1cm] (II1) at (-4.5,-4.2) {\ref{part:algebra}.\ref{sec:clusterCategories} -- Cluster categories};
	\node[ellipse, draw=green, fill=green!10, align=center, text width=4.9cm, inner sep=0, minimum height=1cm] (II2) at (4.5,-4.2) {\ref{part:algebra}.\ref{sec:extricats} -- Extriangulated categories};
 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.180) to [bend right, looseness=1.1] node[label={[shift={(.6,-1.5)}]applied to}]{} (I21.90);
 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.240) to [bend right, looseness=.6] node[label={[shift={(.6,-1.5)}]applied to}]{} (I22.90);
 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=8}, draw=blue, fill=blue!10] (I1.300) to [bend left, looseness=.6] node[label={[shift={(-.6,-1.5)}]applied to}]{} (I23.90);
 	\draw[decorate, decoration={shape backgrounds, shape=circle, shape size=2mm, shape evenly spread=10}, draw=blue, fill=blue!10] (I1.0) to [bend left, looseness=1.1] (I24.90);
 	\draw[decorate, decoration={markings, mark=at position 0.5 with {\draw [red, ultra thick,-] ++ (-.3,-.3)  -- ( .3, .3); \draw [red, ultra thick,-] ++ (-.3, .3)  -- ( .3,-.3);}}] (I1.0) to [bend left, looseness=1.1] node[label={[shift={(-.4,-1.7)}]not applicable to}]{} (I24.90);
	\draw[thick, double distance=2pt, -implies] (I22.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I22.270)+(0,.15);
	\draw[thick, double distance=2pt, -implies] (I23.270)+(0,-1.3) to node[semithick, right, outer sep=0, pos=0.65]{used for} (I23.270)+(0,.15);
	\draw[thick, ->] (II1.0) to node[semithick, above, outer sep=0]{generalizes to} (II2.180);
	\draw let \p{top}=(I1.90), \p{left}=(I21.180), \p{bottom}=(I22.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\draw let \p{top}=([shift={(0,.2)}]II1.90), \p{left}=(I21.180), \p{bottom}=(II1.270), \p{right}=(I24.0) in (\x{left}, \y{top}) -- (\x{left}, \y{bottom}) -- (\x{right}, \y{bottom}) -- (\x{right}, \y{top}) -- (\x{left}, \y{top});
	\path let \p{top}=(I1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-6.15, \y{top}) {\ref{part:geometry} -- Discrete geometry};
	\path let \p{top}=([shift={(0,.2)}]II1.90) in node[rectangle, draw, fill=white, align=left, inner sep=4] (I) at (-5.8, \y{top}) {\ref{part:algebra} -- Representation theory};
\end{tikzpicture}
}
